\chapter{Available Checks}
\label{chap:Available Checks}


    The C style checker provides the following checks:
    
    \begin{description}
        \item[Array Trailing Comma] \hfill \\ 
        Checks if array initialization contains optional trailing comma.
        
        \item[Type Name] \hfill \\
        Checks that type names conform to a format specified by the format property. 

        \item[Default Comes Last] \hfill \\
        Check that the \texttt{default} case is after all the cases in a switch statement. 


         \item[Empty Block] \hfill \\
        Checks for empty blocks.

        \item[Empty Statement] \hfill \\
        Detects empty statements (standalone ';').

        \item[Fall Through] \hfill \\
            Checks for fall through in \texttt{switch} statements Finds locations where a case contains code
            --- but lacks a \texttt{break}, \texttt{return}, or \texttt{continue} statement.

        \item[File Length] \hfill \\
        Checks for long source files.

        \item[File Tab Character] \hfill \\
        Checks to see if a file contains a tab character.

        \item[Generic White space] \hfill \\
        Checks that the white space around the generic tokens $<$ and $>$ are 
        correct to the typical convention.

        \item[Indentation] \hfill \\
        Checks correct indentation of C Code.

        \item[Left Curly] \hfill \\
        Checks the placement of left curly braces on types, functions and other blocks. 

        \item[Line Length] \hfill \\
        Checks for long lines.


        \item[Local Variable Name]  \hfill \\
        Checks that local, variable names conform to a format specified by the format property.

        \item[Magic Number]  \hfill \\
        Checks for magic numbers.


        \item[Method Name] \hfill \\
        Checks that method names conform to a format specified by the format property.

        \item[Method Param Pad] \hfill \\
        Checks the padding between the identifier of a method definition and the left parenthesis of the parameter list.

        \item[Missing Switch Default] \hfill \\
        Checks that switch statement has \texttt{default} case.

        \item[Need Braces] \hfill \\
        Checks for braces around code blocks.

        \item[One Statement Per Line] \hfill \\
        Checks there is only one statement per line.

        \item[Parameter Name] \hfill \\
        Checks that parameter names conform to a format specified by the format property.

         \item[Paren Pad] \hfill \\
        Checks the padding of parentheses; that is whether a space is required after a 
        left parenthesis and before a right parenthesis.

        \item[Right Curly] \hfill \\
        Checks the placement of right curly braces.

        \item[Type Name] \hfill \\
        Checks that type names conform to a format specified by the format property.

        \item[Type cast Paren Pad] \hfill \\
        Checks the padding of parentheses for typecasts.

        \item[Header Comment] \hfill \\
        Detects files missing or improperly formatted header comments.

        \item[White space After] \hfill \\
        Checks that a token is followed by white space.

        \item[White space Around] \hfill \\
        Checks that a token is surrounded by white space.
        
        \item[Multiple variable declarations with initialization] \hfill \\
        Checks that an initialization of a variable is on its own line.

        \item[Function Comment format] \hfill \\
        Checks that functions preceded by a short comment describing the
        function are correctly formatted.


